---
title: "Mausam Duggal"
author: "Use the Border survey to create external trip matrix for GGHM 4.0"
date: "July 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r init, echo=FALSE, message=FALSE}
library(dplyr); library(ggplot2); library(knitr); library(rgeos); library(ggmap); 
library(rgdal); library(sp); library(png); library(spatialEco); library(data.table)
```

Set working directory to start the process
```{r Set Working Directory}
# opts_knit$set(root.dir = 'c:/personal/r')
 wd <- setwd("c:/personal/r")
```

Batch in the Border survey and the **SurveysAll** table in that database. Clean it to get **rid of Weekend** travel. Interestingly enough, trips were only counted from **8:00 a.m. onwards. This could cause some underreporting.** 

```{r batch in the Border survey}

# read the file in its raw format after saving it out as a CSV
bd <- read.csv("BorderSurveys.csv", stringsAsFactors = FALSE)

  # Now only keep the Weekday travel and trips between 600 and 900
  bd_wd <- subset(bd, Weekday.Weekend == "Weekday") %>% subset(., Hour > 5 & Hour < 9)
  bd_wd[is.na(bd_wd)] <- 0
  
```

As one will see, once we get rid of the weekend trips and restrict the records to the a.m. peak period we are left with very few trips.

```{r print the summary before beginning the data processing}

print(paste0("Trips reported in the peak hour are a total of ", sum(bd_wd$ExpFac_hr)))

```


Clean out any unnecssary columns summarise the total trips. One thing to note here is that unlike the Simcoe and Peterborough surveys, if a trip was captured at any of the bridge, it would have to pass through the GGH area. Thus we can eliminate the possibility of external-external trips that might be captured by the border survey and do not cross the GGHM; a possibility we could not ignore in the other two surveys.

```{r only keep relevant columns}
# create column list to keep

keeps <- c("Month", "Day", "Hour", "Direction", "Location", "O1.TAZ", "O2.TAZ", "D1.TAZ", "D2.TAZ", "ExpFac_hr", 
           "O1_Lat", "O1_Long", "D1_Lat", "D1_Long")

bd_wd1 <- bd_wd[keeps]
  bd_wd1[is.na(bd_wd1)] <- 0

#' also bring in the GGHM externals based on the bridge in question
bridge <- c(
  "Rainbow Bridge" = 9502,         
  "Whirlpool Bridge" = 9503,        
  "Peace Bridge" = 9501,         
  "Queenston-Lewiston Bridge" = 9504
)

#' transfer over the GGHM external IDs 
bd_wd2 <- transform(bd_wd1, GGHEx = ifelse(bd_wd1$Location == "Rainbow Bridge", 9502,
                                           ifelse(bd_wd1$Location == "Whirlpool Bridge", 9503,
                                                  ifelse(bd_wd1$Location == "Peace Bridge", 9501, 9504))))

#print the results

print(paste0("The Border Survey points to a total of ", sum(bd_wd1$ExpFac_hr), " external trips, in the morning peak period"))

```

Now add the final O and D for those trips that have one end within the GGHM and the other on one of the bridges

```{r create the final trip ends by GGHM TAZ}

#' first, for trips that have one trip end within the GGHM and the other is one of the bridges
bd_wd2 <- transform(bd_wd2, Origin = ifelse(bd_wd2$Direction == "Into Canada", bd_wd2$GGHEx, 
                                            ifelse(bd_wd2$Direction == "Into U.S.", ifelse(bd_wd2$O2.TAZ > 0, 
                                                                                           bd_wd2$O2.TAZ, bd_wd2$O1.TAZ), 0))) 

bd_wd2 <- transform(bd_wd2, Dest = ifelse(bd_wd2$Direction == "Into U.S.", bd_wd2$GGHEx,
                                            ifelse(bd_wd2$Direction == "Into Canada", ifelse(bd_wd2$D2.TAZ > 0, 
                                                                                             bd_wd2$D2.TAZ, bd_wd2$D1.TAZ), 0)))
                                          
```

Select records that have one trip end outside the GGHM and export it out to geocode in ArcMap and also do a point-to-point assingment

```{r subset records with one trip end out of the GGHM}

#' assign unique IDs to all unique records based on LatLong
bd_wd2 <- transform(bd_wd2, un = paste0(bd_wd2$O1_Lat, bd_wd2$O1_Long, bd_wd2$D1_Lat, bd_wd2$D1_Long))

  # group by lat long and create unique values
  gr <- bd_wd2 %>% group_by(O1_Lat, O1_Long, D1_Lat, D1_Long) %>% summarise(c = n())
    gr <- transform(gr, un = paste0(gr$O1_Lat, gr$O1_Long, gr$D1_Lat, gr$D1_Long)) %>% 
      transform(., id = seq(1, nrow(gr))) %>% subset(., select = c(un, id))

#' join back the ID field
bd_wd2 <- merge(bd_wd2, gr, by.x = "un", by.y = "un", all.x = TRUE)

#' batch out the file for mapping in ArcMap
bd_out <- subset(bd_wd2, Origin == 0 | Dest == 0)
write.csv(bd_out, "border_out.csv")

```

Now batch in the spatially matched files from ArcMap. These shapefiles have been spatially matched to the nearest external GGHM zone.

```{r batch in the files}

#' first batch in the shapefiles that are needed for the spatial joins
orig <- readOGR(wd, "Border_Origin_Ext")
dest <- readOGR(wd, "Border_Dest_Ext")

orig_df <- orig@data %>% subset(., select = c(id, Origin))
  #setnames(orig_df, "ID_1", "NOrig")
dest_df <- dest@data %>% subset(., select = c(id, Dest))
  #setnames(dest_df, "ID_1", "NDest")

```

Now join back the GGHM external ID for each of the missing locations back to the main database

```{r join back the GGHM external information}

# join the GGHM externals
bd_wd3 <- merge(bd_wd2, orig_df, by = "id", all.x = TRUE)
bd_wd3 <- merge(bd_wd3, dest_df, by = "id", all.x = TRUE)

# set all NA to zero
bd_wd3[is.na(bd_wd3)] <- 0

#' now reset all the zones
bd_wd3 <- transform(bd_wd3, Origin = bd_wd3$Origin.x + bd_wd3$Origin.y) %>% 
  transform(., Dest = bd_wd3$Dest.x + bd_wd3$Dest.y)

# sum trips and only keep relevant columns
bd_wd4 <- bd_wd3 %>% group_by(Origin, Dest) %>% summarise(FTrips = sum(ExpFac_hr)) %>% 
  transform(., join = paste0(Origin, Dest))


```

Expand the peak hour surveys to the peak period using the counts observed at the border crossing and as provided by Mauricio on July 15, 2016

```{r expand to peak period}

# multiply by 0.43 to go to peak period

bd_wd4$PP_FTrips <- bd_wd4$FTrips/0.43

write.csv(bd_wd4, "FinalBorder.csv")

```

Report the **results of all the data analysis.**

```{r summary of the results}

print(paste0("The total peak period trips captured at the border stations were ", sum(bd_wd4$PP_FTrips)))

```

