---
title: "Understanding Trip Length Frequency Distributions of Empty Trucks"
output:
  html_notebook: default
  html_document: default
  pdf_document: default
---


```{r}
knitr::opts_chunk$set(echo = TRUE)
```


```{r init, echo=FALSE, message=FALSE}
library(dplyr); library(knitr); library(grid); library(ggplot2); library(gridExtra)

```

#### Understanding Trip Length Frequency Ditributions and Distance Decay Dunction for Empty Trucks

The motivation for calibrating different beta values for medium and heavy trucks and further categoriezed by domestic vs international flows was a result of the availability of the 2012 CVS data and the ability to mine it for empty trucks. It further allowed us to add in a local flavor to the model while asserting the simple but elegant empty truck model framework developed by Rolf and Rick for Chicago.

Function to clean and produce a distance dataframe with the necessary distance bands.
```{r}

clean <- function(truck, distdf){
  "
  Function to bring the observed truck matrix and O-D distance flows
  together and generate a distance matrix that can be used to display
  a trip length frequncy distribution and the cost function

  Arguments: observed truck matrix, and distance by O-D
  Inputs: same as argument
  Retur: distance matrix

  "
  all_obs <- transform(truck, Flag = paste0(Orig,Dest)) %>% 
  subset(., Orig != Dest)

dist <- transform(distdf, Flag = paste0(Orig, Dest))
#' only keep those o-D pairs that match the all_obs truck trip table
dist1 <- subset(dist, dist$Flag %in% all_obs$Flag)

#' add in bands
dist1 <- merge(dist1, all_obs, by = "Flag") %>% 
  arrange(., Dist) %>% 
  transform(., DFlag = ifelse(Dist <= 50, 1, 
                              ifelse(Dist > 50 & Dist <= 100, 2, 
                                     ifelse(Dist > 200 & Dist <= 500, 4, 
                                            ifelse(Dist > 500 & Dist <= 1000, 5, 
                                                   ifelse(Dist > 1000 & Dist <= 2000, 6, 
                                                          ifelse(Dist > 100 & Dist <= 200, 3, 7)))))))

return(dist1)
}

```

Batch in the 2012 observed truck matrices from the CVS data, as produced by Bryce. The data were pre-classified into heavy and medium trucks and by domestic and international flows.
```{r}
#' read observed truck trip matrix and combine
obs_hvy_cb <- read.csv("C:/Projects/Province-wide/Empties/empty_heavy_truck_matrix_int.csv", 
                stringsAsFactors = FALSE)

obs_med_cb <- read.csv("C:/Projects/Province-wide/Empties/empty_straight_truck_matrix_int.csv", 
                    stringsAsFactors = FALSE)

obs_hvy <- read.csv("C:/Projects/Province-wide/Empties/empty_heavy_truck_matrix.csv", 
                stringsAsFactors = FALSE)

obs_med <- read.csv("C:/Projects/Province-wide/Empties/empty_straight_truck_matrix.csv", 
                    stringsAsFactors = FALSE)

dist_cb <- read.csv("C:/Projects/Province-wide/Empties/od_road_distances_all.csv", 
                 stringsAsFactors = FALSE)

dist <- read.csv("C:/Projects/Province-wide/Empties/od_road_distances.csv", 
                 stringsAsFactors = FALSE)
```

Generate the final distance dataframe to be used to understand the TLFD and distance decay function.
```{r}

dist1_cb_hvy <- clean(obs_hvy_cb, dist_cb)
dist1_cb_med <- clean(obs_med_cb, dist_cb)
dist1_hvy <- clean(obs_hvy, dist)
dist1_med <- clean(obs_med, dist)

```

Plot the TLFDs
```{r}
#' Domestic Flows
oh <- ggplot(dist1_hvy, aes(DFlag, Trips)) + geom_bar(stat = "identity", aes(fill = factor(DFlag))) +
  ggtitle("TLFD - Domestic Heavies") + scale_fill_discrete(name="Distance Bands",
                         labels=c("<50km", "50-100", "100-200", "200-500", "500-1000", "1000-2000", ">2000")) + 
  theme(plot.title = element_text(size=10)) + theme(legend.title = element_text(size = 8)) + 
  theme(legend.text = element_text(size = 6))
print("The average distance travelled by a heavy truck for domestic flows is 137.7 kms")
om <- ggplot(dist1_med, aes(DFlag, Trips)) + geom_bar(stat = "identity", aes(fill = factor(DFlag))) +
  ggtitle("TLFD - Domestic Mediums") + scale_fill_discrete(name="Distance Bands",
                         labels=c("<50km", "50-100", "100-200", "200-500", "500-1000", "1000-2000", ">2000")) +
  theme(plot.title = element_text(size=10))  + theme(legend.title = element_text(size = 8)) + 
  theme(legend.text = element_text(size = 6))
print("The average distance travelled by a medium truck for domestic flows is 77.3 kms")
#' International Flows
ch <- ggplot(dist1_cb_hvy, aes(DFlag, Trips)) + geom_bar(stat = "identity", aes(fill = factor(DFlag))) + 
  ggtitle("TLFD - International Heavies") + scale_fill_discrete(name="Distance Bands",
                         labels=c("<50km", "50-100", "100-200", "200-500", "500-1000", "1000-2000", ">2000")) + 
  theme(plot.title = element_text(size=10))  + theme(legend.title = element_text(size = 8)) + 
  theme(legend.text = element_text(size = 6))
print("The average distance travelled by a heavy truck for international flows is 403.9 kms")
cm <- ggplot(dist1_cb_med, aes(DFlag, Trips)) + geom_bar(stat = "identity", aes(fill = factor(DFlag))) +
   ggtitle("TLFD - International Mediums") + scale_fill_discrete(name="Distance Bands",
                         labels=c("<50km", "50-100", "100-200", "200-500", "500-1000", "1000-2000", ">2000")) + 
  theme(plot.title = element_text(size=10))  + theme(legend.title = element_text(size = 8)) + 
  theme(legend.text = element_text(size = 6))
print("The average distance travelled by a medium truck for international flows is 195.9 kms")

grid.arrange(oh, om, ch, cm)

```

Generate the Cost Functions using the calibrated Betas. Different beta values were calculated for heavy and medium trucks for domestic and international travel.
```{r}
#' Domestic Cost FUnctions
d_h <- -0.00332
d_m <- -0.009442

dist1_hvy <- transform(dist1_hvy, Cf = exp(d_h*dist1_hvy$Dist))
dist1_med <- transform(dist1_med, Cf = exp(d_m*dist1_med$Dist))

#' International Const FUnctions
i_h <- -0.006311
i_m <- -0.015388

dist1_cb_hvy <- transform(dist1_cb_hvy, Cf = exp(i_h*dist1_cb_hvy$Dist))
dist1_cb_med <- transform(dist1_cb_med, Cf = exp(i_m*dist1_cb_med$Dist))

```

Plot the cost functions. The **red** lines are for Domestic flows. The **blue** lines are for international flows. The **solid** lines represent Medium trucks, while the **dashed** lines are heavy trucks.
```{r}

p <- ggplot() + 
  geom_line(data = dist1_hvy, aes(x = Dist, y = Cf), color="red", linetype = 2)+geom_point(color="red") + 
  geom_line(data = dist1_med, aes(x = Dist, y = Cf), color="red", linetype = 1)+geom_point(color="red") +
  
  geom_line(data = dist1_cb_hvy, aes(x = Dist, y = Cf), color="blue", linetype = 2)+geom_point(color="blue") + 
  geom_line(data = dist1_cb_med, aes(x = Dist, y = Cf), color="blue", linetype = 1)+geom_point(color="blue") +
  ggtitle("Cost Functions for Domestic and International Flows categorized by Medium and Heavy Trucks")
p

```


**Heavy** empty trucks are the most sensitive to distance, for **domestic** flows. 

**Medium** trucks are more sensitive than heavy trucks to distance.


Of the **Medium** empty trucks, **international** flows are more sensitive to distance than 
**domestic** flows. 
