---
title: "External Canadian Emp by NAICS and Population by Province"
author: "Mausam Duggal"
date: "August 6, 2016"
output: html_document
---

```{r init, echo=FALSE, message=FALSE}
library(dplyr); library(knitr); library(gaussfacts); library(rmsfact); library(reshape2); library(rgdal); library(sp); library(stringr)

wd <- setwd("c:/personal/r")
```

#### Batch in the 2011 NHS Data 

Use the 2011 NHS semi-custom tabulations to calculate the employment by NAICS for provinces other than Manitoba and Quebec.

```{r NHS}
nhs <- read.csv("c:/personal/r/2011NHS.csv", stringsAsFactors = FALSE)

```

Only select the columns that contain employment information and then filter rows for the various provinces. Upon which create new column names to coresspond to NAICS heading and add in a placeholder for the provincial external id.

```{r subset data}

#' select columns that coresspond to employment data by NAICS
nhs1 <- nhs[c(1:2, 2066:2088)]

#' create list of names to subset 
names = c('Newfoundland', 'Nova', 'Saskatch', 'Alberta', 'New Brunswick', 'Prince Edward Island', 'Nunavut', 'Northwest Territories', 'Yukon', 'British Columbia')

# select only rows that coresspond to the names and then get rid of unncessary rows that come in as a result of the names list
nhs3 <- nhs1[grep(paste0(names, collapse = '|'), nhs1$Geography, ignore.case = TRUE),]
nhs4 <- nhs3[-c(5,6,8,9,11,12,15,16),] %>% .[-c(3, 5)]

nhs4$Geography <- gsub("/.*", "", nhs4$Geography) 
  nhs4$Geography <- str_replace(nhs4$Geography, "\\(.*\\)", "")
```

```{r columnnames}
# create list of colnames
cols = c("GeogCMA", "Population", "NAICS_99", "NAICS_11", "NAICS_21", "NAICS_22", "NAICS_23", "NAICS_31", "NAICS_41", "NAICS_44", "NAICS_48", "NAICS_51", "NAICS_52", "NAICS_53", "NAICS_54", "NAICS_55", "NAICS_56", "NAICS_61", "NAICS_62", "NAICS_71", "NAICS_72", "NAICS_81", "NAICS_91" )

```

```{r}

# set column names and add column for external ID
colnames(nhs4) <- cols
nhs4 <- transform(nhs4, ExtID = 0)

```

#### Batch in the DA shapefile for Manitoba and Quebec

The non-CMA DAs have a CMAUID of 999 which will be used in conjunction with the rest of the CMAUIDs to generate emp amd pop estimates

```{r}
da <- readOGR(".", "QM_DA")

```

```{r get database table}
da.df <- da@data 

# only keep certain columns
keep = c("DAUID", "CMAUID")
da.df <- da.df[keep]
  # get rid of factors or it won't join the dataframes
  da.df$CMAUID <- as.numeric(as.character(da.df$CMAUID))
   da.df$DAUID <- as.numeric(as.character(da.df$DAUID))

```


```{r transfer CMAUID to NHS}

#' create list of names to subset 
names1 = c('46', '24')

# select only rows that coresspond to the names1 and then get rid of text in parenthesis for joining with shapefile data
nhs5 <- subset(nhs1, grepl("^(46|24)", nhs1$Geography))

nhs5$Geography <- str_replace(nhs5$Geography, "\\(.*\\)", "") 
  nhs5$Geography <- as.numeric(as.character(nhs5$Geography))

# now join the data 
colnames(nhs5)[1] <- "DAUID"
nhs6 <- merge(nhs5, da.df, by.x = "DAUID", by.y = "DAUID", all.x = TRUE)

```

```{r}
# drop unncessary columns
nhs6 <- nhs6[-c(1,3,5)]
```

```{r}
# column names
cols1 = c("Population", "NAICS_99", "NAICS_11", "NAICS_21", "NAICS_22", "NAICS_23", "NAICS_31", "NAICS_41", "NAICS_44", "NAICS_48", "NAICS_51", "NAICS_52", "NAICS_53", "NAICS_54", "NAICS_55", "NAICS_56", "NAICS_61", "NAICS_62", "NAICS_71", "NAICS_72", "NAICS_81", "NAICS_91", "GeogCMA" )

```

```{r}
colnames(nhs6) <- cols1
nhs6 <- transform(nhs6, ExtID = 0)

```

```{r}

nhs6.sum <- nhs6 %>% group_by(GeogCMA) %>% summarise_each(funs(sum))

```

#### Bring in the external IDs and bind the dataframes together

```{r}
# file with provincial IDs
equi <- read.csv("C:/personal/r/extid_equiv.csv", stringsAsFactors = FALSE)

```


```{r}

final <- rbind(nhs4, nhs6.sum)

```

```{r}

final1 <- merge(final, equi, by.x = "GeogCMA", by.y = "GeogCMA", all.x = TRUE) 

# create equivalency for provinces
prov <- c(
  "Alberta" = 9793,
  "British Columbia" = 9794,
  "New Brunswick" = 9791,
  "Newfoundland and Labrador" = 9790,
  "Northwest Territories" = 9796,
  "Nova Scotia" = 9787,
  "Nunavut" = 9795,
  "Prince Edward Island" = 9789,
  "Saskatchewan" = 9792,
  "Yukon" = 9797
)

final1[39:48,]$ProvID <- prov[1:10] 
final1 <- subset(final1, select = - ExtID)

```

```{r}

write.csv(final1, "extsociodata.csv")

```


