---
title: "Make and Use Coefficients"
author: "Mausam Duggal"
date: "September 6, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r init, echo=FALSE, message=FALSE}
library(dplyr); library(knitr); library(reshape2); library(ggplot2); library(png); library(grid)

wd <- setwd("c:/personal/r")
```

### Life over the last 2 weeks as we realize why **freight modellers** lose all their hair

```{r, fig.width = 12, fig.height = 8}

img <- readPNG("emotions.png") 
      grid.raster(img)

```


### Calculate the Make and Use coefficients for the Commodity Flow Model Using StatsCan. 


Bryce spoke to Chris Bachmann and he forwarded a StatsCAN link that has the make and use tables. Bryce downloaded the entire table, which is being batched in here.  [Detailed tables from statscan](http://www5.statcan.gc.ca/cansim/a26?lang=eng&retrLang=eng&id=3810033&&pattern=&stByVal=1&p1=1&p2=-1&tabMode=dataTable&csid=)

```{r I am sick and tired of NAs so a function to set them for good}

f_rep <- function(df) {
  # this function is used to set all NA values to zero in a dataframe
df[is.na(df)] <- 0
return(df)
}

```


```{r read Oregon MakeUse file}
mu <- read.csv("c:/personal/r/I-O_Structure_CAD_Economy.csv", stringsAsFactors = FALSE)    
```

#### Make the two tables.

The tables contains many categories in the commodity and industry category that do not corresspond to a MAKE and USE paradigm. Excluding those categories, both the INDUSTRY and PRODUCT fields have around 400 unique values. The unique values in each were batched out and the corressponding SCTG2 and NAICS2 codes were manually entered. These unique equivalency files are being batched below. 

```{r}
#' batch in the unique NAICS and Commodity equivalencies
naics <- read.csv("c:/personal/r/IndustryCodes.csv", stringsAsFactors = FALSE) 
comm <- read.csv("c:/personal/r/CommodityCodes.csv", stringsAsFactors = FALSE) %>% f_rep(.)

```

```{r merge the datasets}

mu1 <- merge(mu, naics, by.x = 'INDUSTRY', by.y = 'Ind_Desc', all.x = TRUE) %>% 
  merge(., comm, by.x = 'PRODUCT', by.y = 'Comm_Desc', all.x = TRUE) %>% na.omit(.)

```


```{r MAKE TABLE}
# Group by NAICS and SCTG to create the make table
make <- mu1 %>% group_by(NAICS2, SCTG2) %>% summarise(Tval = sum(Value))

# cast into wide format
make_wide <- dcast(make, NAICS2 ~ SCTG2, value.var = "Tval") %>% f_rep(.)

# get column names and then set them in the dataset
make_wide1 <- make_wide[, -1]
rownames(make_wide1) <- make_wide[, 1]

#' add in the SCTG2 columns that are presented as a range
make_wide1$'9' <- make_wide1$`8`
make_wide1$'11' <- make_wide1$`10`
make_wide1$'12' <- make_wide1$`10`
make_wide1$'17' <- make_wide1$`16`
make_wide1$'18' <- make_wide1$`16`

#' Add total column 
make_wide1$Total <- rowSums(make_wide1)

# get proportions/coefficients
make_wide2 <- make_wide1/make_wide1$Total 

# write out MAKE file
write.csv(make_wide2, "MakeTable.csv")

```


```{r USE TABLE}
# Group by NAICS and SCTG to create the make table
use <- mu1 %>% subset(., SUTABLE = "Use") %>% group_by(SCTG2, NAICS2) %>% summarise(Tval = sum(Value))
# cast into wide format
use_wide <- dcast(use, SCTG2 ~ NAICS2, value.var = "Tval") %>% f_rep(.)

# get column names and then set them in the dataset
use_wide1 <- use_wide[, -1]
rownames(use_wide1) <- use_wide[, 1]

#' add in the SCTG2 rows that are presented as a range
use_wide1['9',]<- use_wide1['8',]
use_wide1['11',]<- use_wide1['10',]
use_wide1['12',]<- use_wide1['10',]
use_wide1['17',]<- use_wide1['16',]
use_wide1['18',]<- use_wide1['16',]

#' Add total column 
use_wide1$Total <- rowSums(use_wide1)

# get proportions/coefficients
use_wide2 <- use_wide1/use_wide1$Total 

# write out USE file
write.csv(use_wide2, "UseTable.csv")

```

### Plot the tables

```{r}
# plot the Make table.
make_temp <- make_wide2[, 1:42]
make_temp$naics2 <- rownames(make_wide2)

make_melt <- melt(make_temp, id.vars = "naics2") %>% setNames(., c("naics2", "sctg2", "coeff")) 
  make_melt$sctg2 <- as.numeric(as.character(make_melt$sctg2))

ggplot(make_melt, aes(y = coeff, x = factor(naics2), fill = factor(sctg2))) + geom_bar(stat = "identity") + coord_flip() + ggtitle("Make Table") + 
  annotate("text", x = c(24, 22, 20, 18, 16, 14, 12, 10, 8, 6), y = 1.10, label = c("sctg 1:Animals and Food...Alcoholic products", "sctg 10-12: Sands", "sctg 13-24: Minerals...Plastics and Rubber", "sctg 25-30: Logs...textiles and leather", "sctg 31: Non-metallic mineral", "sctg 32-34: Base metals...machinery", "sctg 35: Electronic", "sctg 36-38: Motorized...precision instruments", "sctg 39: Furniture", "sctg40-43: Miscellaneous...Mixed freight"), size = 2)

```

```{r}
# plot the USE table.
use_temp <- use_wide2[, 1:24]
use_temp$sctg2 <- rownames(use_wide2)

use_melt <- melt(use_temp, id.vars = "sctg2") %>% setNames(., c("sctg2", "naics2", "coeff")) 
  use_melt$naics2 <- as.numeric(as.character(make_melt$naics2))

ggplot(use_melt, aes(y = coeff, x = factor(sctg2), fill = factor(naics2))) + geom_bar(stat = "identity") + coord_flip() + ggtitle("Use Table") + 
  annotate("text", x = c(24, 22, 20, 18, 16, 14, 12, 10, 8, 6), y = 1.10, label = c("sctg 1:Animals and Food...Alcoholic products", "sctg 10-12: Sands", "sctg 13-24: Minerals...Plastics and Rubber", "sctg 25-30: Logs...textiles and leather", "sctg 31: Non-metallic mineral", "sctg 32-34: Base metals...machinery", "sctg 35: Electronic", "sctg 36-38: Motorized...precision instruments", "sctg 39: Furniture", "sctg40-43: Miscellaneous...Mixed freight"), size = 2.5)

```
